some title

### 2018-02-03 2018/chapters/2_regression.tex
restructuring finished, fwl v2

\begin{figure}[h!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/tss_sum_base.pdf}
\label{fig:tss_base} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/tss_sum_rss.pdf}
\label{fig:tss_rss}}
\caption{\subref{fig:tss_base}: Vectors $y \in \mathbb{R}^{n-2}$ and $\hat y \in Lin(x, \mathbf{1})$; \subref{fig:tss_rss}: Residual sum of squares.}
\end{center}
\end{figure}

@@

\begin{figure}[h!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/tss_sum_tss.pdf}
\label{fig:tss_tss}}
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/tss_sum_ess.pdf}
\label{fig:tss_ess}}
\caption{\subref{fig:tss_tss}: Total sum of squares; \subref{fig:tss_ess}: Explained sum of squares.}
\end{center}
\end{figure}

@@

\begin{figure}[h!]
\center{\includegraphics{figures/tss_sum_final.pdf}}
\caption{$(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2$}
\label{fig:tss_final}
\end{figure}

@@

Consider Figure~\ref{fig:tss_final} from the previous proof.
It was shown there that the vectors $y - \bar y$, $y - \hat y$ and $\hat y - \bar y$ form a right triangle.
Having defined the determination coefficient as
\[
R^2 = \frac{ESS}{TSS}
\]
we conclude that its geometric interpretaion is
\[
\sqrt{R^2} = \frac{\sqrt{ESS}}{\sqrt{TSS}} = \cos \varphi
\]
as shown in Figure~\ref{fig:r_sq_angle}.

\begin{figure}
  \center{\includegraphics{figures/r_sq_ss.pdf}}
  \caption{Determination coefficient as squared $\cos \varphi$}
  \label{fig:r_sq_angle}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/averages_1.pdf}
\label{fig:averages_lin} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/averages_2.pdf}
\label{fig:averages_bars}}
\caption{\subref{fig:averages_lin}: Regression of $y$ on $Lin(\mathbb{1},x)$; \subref{fig:averages_bars}: Regression of $y$ and $\hat y$ on $\mathbb{1}$.}
\end{center}
\end{figure}

@@

\begin{theorem}
Consider regression
\begin{equation} \label{eq:fwl_1}
y = X_1 \beta_1 + X_2 \beta_2 + u
\end{equation}
where $X_{n \times k} = [X_1 X_2]$, i.e. $X_1$ consists of first $k_1$ columns of $X$ and $X_2$ consists of remaining $k_2$ columns of $X$,
$\beta_1$ and $\beta_2$ are comfortable, i.e. $k_1 \times 1$ and $k_2 \times 1$ vectors.
Consider another regresison
\begin{equation}  \label{eq:fwl_2}
M_1 y = M_1 X_2 \beta_2 + M_1 u
\end{equation}
where $M_1 = I - P_1$ projects onto the orthogonal complement of the column space of $X_1$ and $P_1 = X_1(X_1^TX_1)^{-1}X_1^T$ is the projection onto the column space of $X_1$.
Then the estimate of $\beta_2$ from regression~\ref{eq:fwl_1} will be the same as the estimate from regression~\ref{eq:fwl_2}.
\end{theorem}

\begin{proof}
Geometrical proof will be presented for the following model:
\begin{equation} \label{eq:fwl_proof}
y_i = \beta_1 x_i + \beta_2 z_i + u_i
\end{equation}

We start with regression `all-at-once' and will distinct its coefficients with index $(1)$. The only step in obtaining $\beta_1^{(1)}$ is regressing $y$ on $Lin(x,z)$ and then expanding $\hat y$ as a linear combination of basis vectors $x$ and $z$,
which is shown in Figure~\ref{fig:fwl_1_regression_3d}. Figure~\ref{fig:fwl_1_regression_lin} depicts $Lin(x, z)$.

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/fwl_1_regression.pdf}
%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
\label{fig:fwl_1_regression_3d} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/fwl_1_regression_lin.pdf}
%\caption{$Lin(x, z)$}
\label{fig:fwl_1_regression_lin}}
\caption{\subref{fig:fwl_1_regression_3d}: Regression of $y$ on $Lin(x,z)$; \subref{fig:fwl_1_regression_lin}: $Lin(x, z)$.}
\end{center}
\end{figure}

As for the model~\ref{eq:fwl_2}, where several regressions are performed consecutively, we start with regressing $y$ on $z$, resulting in $\tilde{y}$, which we will refer to as ``cleansed'' $y$.

@@

Following that, $x$ is regressed on $z$, resulting in $\tilde{x}$ — ``cleansed'' $x$.

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/fwl_2_regression.pdf}
%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
\label{fig:fwl_2_regression_first} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/fwl_2_regression_translation.pdf}
%\caption{$Lin(x, z)$}
\label{fig:fwl_2_regression_trans}}
\caption{\subref{fig:fwl_2_regression_first}: Regression of $y$ on $z$ and of $x$ on $z$;
\subref{fig:fwl_2_regression_trans}: Translation of $\tilde{x}$.}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/fwl_3_3d.pdf}
%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
\label{fig:fwl_3_3d} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/fwl_3_lin.pdf}
%\caption{$Lin(x, z)$}
\label{fig:fwl_3_lin}}
\caption{\subref{fig:fwl_3_3d}: Point A stands for the origin, B — $\hat\gamma z$,
C — $x$, D — $\hat\alpha z$, E — intersection of vector $x$ and line parallel to $\tilde x$,
F — $\hat\beta_1^{(1)} x$, G — $\hat\beta_1^{(2)} \tilde{x}$; \subref{fig:fwl_3_lin}: $Lin(x,z)$.}
\end{center}
\end{figure}

In Figure~\ref{fig:fwl_3_lin} segments AF and BH = DG stand for $\hat\beta_1^{(1)}x$
and $\hat\beta_1^{(2)}\tilde x$ respectively, while segments AC and BC represent $x$ and $\tilde{x}$.
Having two congruent angles, triangles ABC and FHC are simillar.
Then, it follows:
\[
\frac{AF}{AC} = \frac{BH}{BC} \Leftrightarrow \frac{\hat\beta_1^{(1)}x}{x} = \frac{\hat\beta_1^{(2)}\tilde x}{\tilde x} \Leftrightarrow \hat\beta_1^{(1)} = \hat\beta_1^{(2)}
\]
\end{proof}

>>>>

\begin{figure}[h!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_rss_ess_tss_yhat.pdf}
\label{fig:tss_base} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_rss_ess_tss_sqr_rss.pdf}
\label{fig:tss_rss}}
\caption{\subref{fig:tss_base}: Vectors $y \in \mathbb{R}^{n-2}$ and $\hat y \in Lin(x, \mathbf{1})$; \subref{fig:tss_rss}: Residual sum of squares.}
\end{center}
\end{figure}

@@

\begin{figure}[h!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_rss_ess_tss_sqr_tss.pdf}
\label{fig:tss_tss}}
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_rss_ess_tss_sqr_ess.pdf}
\label{fig:tss_ess}}
\caption{\subref{fig:tss_tss}: Total sum of squares; \subref{fig:tss_ess}: Explained sum of squares.}
\end{center}
\end{figure}

@@

\begin{figure}[h!]
\center{\includegraphics{figures/02_rss_ess_tss_final.pdf}}
\caption{$(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2$}
\label{fig:tss_final}
\end{figure}

@@

Consider Figure~\ref{fig:tss_final} from the previous proof.
It was shown there that the vectors $y - \bar y$, $y - \hat y$ and $\hat y - \bar y$ form a right triangle.
Having defined the determination coefficient as
\[
R^2 = \frac{ESS}{TSS}
\]
we conclude that its geometric interpretaion is
\[
R^2 = \frac{ESS}{TSS} = \cos^2 \varphi
\]
as shown in Figure~\ref{fig:r_sq_angle}.

\begin{figure}
  \center{\includegraphics{figures/02_determination_coefficient.pdf}}
  \caption{Determination coefficient as squared $\cos \varphi$}
  \label{fig:r_sq_angle}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_averages_yhat_decomposed.pdf}
\label{fig:averages_lin} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_averages_final.pdf}
\label{fig:averages_bars}}
\caption{\subref{fig:averages_lin}: Regression of $y$ on $Lin(\mathbf{1},x)$; \subref{fig:averages_bars}: Regression of $y$ and $\hat y$ on $\mathbf{1}$.}
\end{center}
\end{figure}

@@

\begin{theorem}
Consider regression
\begin{equation} \label{eq:fwl_1}
y = X_1 \beta_1 + X_2 \beta_2 + u
\end{equation}
where $X_{n \times k} = [X_1 X_2]$, i.e. $X_1$ consists of first $k_1$ columns of $X$ and $X_2$ consists of remaining $k_2$ columns of $X$,
$\beta_1$ and $\beta_2$ are comfortable, i.e. $k_1 \times 1$ and $k_2 \times 1$ vectors.
Consider another regresison
\begin{equation}  \label{eq:fwl_2}
M_1 y = M_1 X_2 \beta_2 + M_1 u
\end{equation}
where $M_1 = I - P_1$ projects onto the~orthogonal complement of the~column space~of
$X_1$ and $P_1 = X_1(X_1^TX_1)^{-1}X_1^T$ is the~projection onto the~column space of~$X_1$.
Then the~estimate of $\beta_2$ from regression~\ref{eq:fwl_1} will be the~same
as the~estimate from regression~\ref{eq:fwl_2}.
\end{theorem}

There are two ways to~visualize the~proof of~the~Frisch-Waugh-Lovell theorem
using geometric concepts. Both are presented below.

\begin{proof}
1.~Consider the following model:
\begin{equation} \label{eq:fwl_proof}
y_i = \beta_1 x_i + \beta_2 z_i + u_i
\end{equation}

We start with regression `all-at-once' and will distinct its coefficients with~index $(1)$.
The~only step in~obtaining $\beta_1^{(1)}$ is regressing $y$ on~$Lin(x,z)$ and
then expanding $\hat y$ as a~linear combination of~basis vectors $x$ and $z$,
which is shown in~Figure~\ref{fig:fwl_1_regression_3d}. Figure~\ref{fig:fwl_1_regression_lin}
depicts $Lin(x, z)$.

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_yhat_decomposed.pdf}
\label{fig:fwl_1_regression_3d} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_yhat_decomposed_lin.pdf}
\label{fig:fwl_1_regression_lin}}
\caption{\subref{fig:fwl_1_regression_3d}: Regression of $y$ on $Lin(x,z)$; \subref{fig:fwl_1_regression_lin}: $Lin(x, z)$.}
\end{center}
\end{figure}

As for the~model~\ref{eq:fwl_2}, where several regressions are performed consecutively,
we start with regressing $y$~on~$z$, resulting in $\tilde{y}$,
which we will refer~to as~`cleansed' $y$.

@@

Following that, $x$ is regressed on $z$, resulting in $\tilde{x}$ — `cleansed' $x$.

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_cleansed_variables.pdf}
\label{fig:fwl_2_regression_first} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_translation.pdf}
\label{fig:fwl_2_regression_trans}}
\caption{\subref{fig:fwl_2_regression_first}: Regression of $y$ on $z$ and of $x$ on $z$;
\subref{fig:fwl_2_regression_trans}: Translation of $\tilde{x}$.}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_final.pdf}
\label{fig:fwl_3_3d} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_final_lin.pdf}
\label{fig:fwl_3_lin}}
\caption{\subref{fig:fwl_3_3d}: Point A stands for the origin, B — $\hat\gamma z$,
C — $x$, D — $\hat\alpha z$, E — intersection of vector $x$ and line parallel to $\tilde x$,
F — $\hat\beta_1^{(1)} x$, G — $\hat\beta_1^{(2)} \tilde{x}$; \subref{fig:fwl_3_lin}: $Lin(x,z)$.}
\end{center}
\end{figure}

In Figure~\ref{fig:fwl_3_lin} segments $AF$ and $BH = DG$ stand for $\hat\beta_1^{(1)}x$
and $\hat\beta_1^{(2)}\tilde x$ respectively, while segments AC and BC represent $x$ and $\tilde{x}$.
Having two congruent angles, triangles ABC and FHC are simillar.
Then, it follows:
\[
\frac{AF}{AC} = \frac{BH}{BC} \Leftrightarrow \frac{\hat\beta_1^{(1)}x}{x} = \frac{\hat\beta_1^{(2)}\tilde x}{\tilde x} \Leftrightarrow \hat\beta_1^{(1)} = \hat\beta_1^{(2)}
\]

2. Alternatively, we could implement a~concept close to the~partial correlation.
In the~same model~\ref{eq:fwl_proof} we wiil treat $z$ vector fixed and again
consecutively cleanse the $x$ and $y$ variables by projecting them onto
the~space orthogonal to $z$, i.e., $Lin^\perp(z)$ as demonstrated in Figure~\ref{fig:fwl_v2_rcleansed}.
Then we perform a~regression of the~`cleansed' $\tilde y$ on the~`cleansed' $\tilde x$
(see Figure\ref{fig:fwl_v2_regression_cleansed}).

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/02_fwl_v2_cleansed_variables.pdf}
\label{fig:fwl_v2_rcleansed}}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/02_fwl_v2_cleansed_regression.pdf}
\label{fig:fwl_v2_regression_cleansed}}
\caption{\subref{fig:fwl_2_regression_first}: `Cleansed' variables $\tilde x$ and $\tilde y$;
\subref{fig:fwl_2_regression_trans}: `Cleansed' $\tilde y$ regressed on `cleansed' $\tilde{x}$.}
\end{center}
\end{figure}

Now we show that the~latter regression produces $\hat \beta_1$ coefficient which
is exatly the coefficient from the~`one-step' regression of~original $y$ onto
original $x$ and $z$. Recall that the~vector $y$ can be split~up into a~sum of
some multiple of $x$ and some multiple of $z$. Since the second term is
the~orthogonal component its projection yields zero. The~multiple of $x$
is equal to $\hat \beta_1$ by construction.

Assume that the coefficient at $\tilde x$ is some unknown variable $\hat \delta$.
Then consider the similar triangles in the $Lin(x,z)$. From the~proportions
we obtain:
\[
\frac{CE}{CA} = \frac{CD}{CB} \Leftrightarrow \frac{\hat \beta_1 x}{x} = \frac{\hat \delta \tilde x}{\tilde x} \Rightarrow \hat \beta_1 = \hat \delta
\]

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/02_fwl_v2_similar_triangles.pdf}
\label{fig:fwl_v2_triangles}}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/02_fwl_v2_final.pdf}
\label{fig:fwl_v2_final}}
\caption{\subref{fig:fwl_2_regression_first}: Similar triangles: $\bigtriangleup ABC \sim \bigtriangleup EDC$;
\subref{fig:fwl_2_regression_trans}: Alternative proof for the Frisch-Waugh-Lovell theorem.}
\end{center}
\end{figure}

### 2018-02-03 2018/chapters/3_partial_correlation.tex
restructuring finished, fwl v2

\begin{figure}
  \centering
  \includegraphics[width=0.55\linewidth]{figures/partial_corr.pdf}
  \caption{Partial correlation between $X$ and $Y$ while $Z$ is fixed.}
  \label{fig:pcorr_def1}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.55\linewidth]{figures/part_corr_2.pdf}
\label{fig:part_corr_alt}}
%\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.35\linewidth]{figures/part_corr_2_lin.pdf}
\label{fig:part_corr_alt_lin}}
\caption{\subref{fig:part_corr_alt}: Alternative definition of the partial correlation;
\subref{fig:part_corr_alt_lin}: $Lin^{\perp}(Z)$.}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/pcorr_theorem_x.pdf}
\label{fig:pcorr_t_x}}
%\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/pcorr_theorem_y.pdf}
\label{fig:pcorr_t_y}}
\caption{\subref{fig:pcorr_t_x}: $\hat u$ form regression of x onto $Y$ and $Z$, $\hat u$ projected;
\subref{fig:pcorr_t_y}: $\hat v$ from regression of $Y$ onto $X$ and $Z$, $\hat v$ projected.}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/pcorr_t_xy.pdf}
\label{fig:pcorr_t_xy}}
%\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/pcorr_t_lin.pdf}
\label{fig:pcorr_t_lin}}
\caption{\subref{fig:pcorr_t_x}: The residuals of both regressions;
\subref{fig:pcorr_t_y}: $Lin^{\perp}(Z)$.}
\end{center}
\end{figure}
\end{proof}

>>>>

\begin{figure}
  \centering
  \includegraphics[width=0.55\linewidth]{figures/03_partial_correlation_definition.pdf}
  \caption{Partial correlation between $X$ and $Y$ while $Z$ is fixed.}
  \label{fig:pcorr_def1}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.55\linewidth]{figures/03_partial_correlation_regression_definition.pdf}
\label{fig:part_corr_alt}}
%\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.35\linewidth]{figures/03_partial_correlation_regression_definition_lin.pdf}
\label{fig:part_corr_alt_lin}}
\caption{\subref{fig:part_corr_alt}: Alternative definition of the partial correlation;
\subref{fig:part_corr_alt_lin}: $Lin^{\perp}(Z)$.}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/03_partial_correlation_residuals_x.pdf}
\label{fig:pcorr_t_x}}
%\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/03_partial_correlation_residuals_y.pdf}
\label{fig:pcorr_t_y}}
\caption{\subref{fig:pcorr_t_x}: $\hat u$ form regression of x onto $Y$ and $Z$, $\hat u$ projected;
\subref{fig:pcorr_t_y}: $\hat v$ from regression of $Y$ onto $X$ and $Z$, $\hat v$ projected.}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/03_partial_correlation_residuals_xy.pdf}
\label{fig:pcorr_t_xy}}
%\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{figures/03_partial_correlation_residuals_xy_lin.pdf}
\label{fig:pcorr_t_lin}}
\caption{\subref{fig:pcorr_t_x}: The residuals of both regressions;
\subref{fig:pcorr_t_y}: $Lin^{\perp}(Z)$.}
\end{center}
\end{figure}
\end{proof}
